---
title: "Shelter Journal"
date: "`r Sys.Date()`"
author: "[Fran Pérez](https://es.linkedin.com/in/franjperez)"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: journal
    highlight: tango
---

<style type="text/css">
div.sourceCode + table.table {
    width: 100%;
}

table.table {
    width: auto;
    margin: 20px auto;
    background-color: white;
}

.section.level5.collapse {
	padding: 1px 4px;
	background-color: #FBE8E7;
}

pre.r { 
    margin:0; 
}

pre.r  + pre {
    border-top: 0;
    margin: 0;
}

pre + p {
    margin-top: 10px;
}

blockquote {
    font-size: 16.5px !important;
}

div.sourceCode + p {
    margin-top: 10px;
}

pre.sourceCode.r {
    font-size: 13px;
}

p img {
    display: block;
    margin: 2px auto;
}

table + .caption {
    text-align: center;
    margin-top: -24px;
    padding: 10px;
    font-style: italic;
}

.glyphicon-eye-open {
    color: #eb6864;
    padding: 17px;
}
</style>

## Introduction

[Shelter animal outcomes](https://www.kaggle.com/c/shelter-animal-outcomes)

[Quick & Dirty RandomForest](https://www.kaggle.com/mrisdal/shelter-animal-outcomes/quick-dirty-randomforest)

[XGB and 'Quick and Dirty' Ensemble](https://www.kaggle.com/jwjohnson314/shelter-animal-outcomes/xgb-and-quick-and-dirty-ensemble)

[dog groups](https://www.kaggle.com/andraszsom/shelter-animal-outcomes/dog-breeds-dog-groups)

[Uncertainty estimates of outcome types](https://www.kaggle.com/andraszsom/shelter-animal-outcomes/uncertainty-estimates-of-outcome-types)


## Feature engineering

Loading packages used for feature engineering

```{r warning=F,error=F,message=F}
library(readr) # csv file read / write
library(ggplot2) # visualization
library(ggthemes) # visualization
library(dplyr) # data manipulation
library(lubridate) # dates
library(caret) # classification algorithm
library(stringr)
library(mice) # imputation
``` 

```{r, include=F}
options(width = 120)
options(error = function() {
    save.image(file = "error.rmd.RData")
})
knitr::opts_chunk$set(fig.align = 'center')
set.seed(3874)
``` 

I used `read_csv`function in order to read training and test shelter data

```{r}
csv.folder <- ""
#kaggle folder
#csv.folder <- "../input/"
train <- read_csv(file = paste0(csv.folder, "train.csv"))
test <- read_csv(file = paste0(csv.folder, "test.csv"))
``` 

Then, I combine both data frames into a single one, named `full`. 
I did some column renaming in order to normalize naming between datasets.

```{r}
full <- train %>% 
	rename(Id = AnimalID) %>% 
	select(-OutcomeSubtype) %>% 
	bind_rows(test %>% 
		rename(Id = ID) %>% 
		mutate(OutcomeType = NA, Id = as.character(Id))
	)

full.original <- full
rows.train <- !is.na(full$OutcomeType)

glimpse(full, width=100)
``` 
The full dataset contains 9 variables and 38185 rows of data

| Column     | Description    |
|------------|----------------|
| Id | animal Id |
| Name | animal name (could be nameless)|
| DataTime | timestamp |
| OutcomeType | animal outcome |
| AnimalType | dog or cat |
| SexuponOutcome | animal sex upon outcome |
| AgeuponOutcome | animal age upon outcome |
| Breed | animal's breed |
| Color | animal fur color |

There were another columns in the original files, but were not relevant for analysis.

### Breed

```{r}
sample(unique(full$Breed), 30)
``` 

There is clearly two kinds of breeds: pure and mixed, so let's flag mixed
```{r}
full$Breed.IsMix <- grepl("Mix$", full$Breed)
```

On a second inspection, breed feature seem to have a hair feature embedded. Let's examine it using regular expressions

```{r}
unique(str_match(full$Breed,"(?i)[ a-z]*hair[a-z]*")[,1])
``` 

I count 5 different hair categories: long, medium, short, wire, hairless.

```{r}
# take first part of breed
breed.main <- sapply(full$Breed, function(x) gsub(' Mix', '', strsplit(x, split = '/')[[1]][1]))
# then remove hair
breed.main <- str_trim(gsub("Longhair|Shorthair|Wirehaired|Wire Hair|Wirehair|Hairless|Medium Hair", "", breed.main))

full$Breed.Main <- breed.main
``` 

Let's create hair feature

```{r}
full$Breed.Hair <-	ifelse (grepl("Longhair", full$Breed), "longhair",
					ifelse (grepl("Shorthair", full$Breed), "shorthair",
					ifelse (grepl("Wirehaired|Wire Hair|Wirehair", full$Breed), "wirehair",
					ifelse (grepl("Hairless", full$Breed), "hairless",
					ifelse (grepl("Medium Hair", full$Breed), "medium", "unknown")))))

table(full$Breed.Hair, full$AnimalType)
``` 

Finally, we can add the size ... for the moment, just for dogs and for pure breed

```{r eval=F}
#dogs
dogSizes <- read_csv("dog_size.csv")

full[full$IsMix==F,] <- right_join (dogSizes[,c("Breed.Main","Size")], full[full$IsMix==F,])
full$Size[is.na(full$Size)] <- "unknown"

#cats
#catBreed <- read_csv("cat_breed.csv")
#full[full$IsMix==F,] <- right_join (catBreed[,c("Breed.Main","Size")], full[full$IsMix==F,])
``` 


### Name

Does having a name is correlated with the outcome?

```{r}
full$Name.NoName <- is.na(full$Name)
prop.table(table(full$Name.NoName))
``` 

### Date

We can extract several features from outcome date. 
This features are important because they can highlight seasonal behaviors.

```{r}
full$Hour <- hour(full$DateTime)
full$Weekday <- wday(full$DateTime, label = T)
full$Month <- month(full$DateTime, label = T)
full$Year <- year(full$DateTime)

timeOfDay <- ifelse(full$Hour > 5 & full$Hour < 11, 'morning',
                  ifelse(full$Hour > 10 & full$Hour < 16, 'midday',
                  ifelse(full$Hour > 15 & full$Hour < 20, 'lateday', 'night')))

full$TimeOfDay <- factor(timeOfDay, levels = c("morning", "midday", "lateday", "night"))
rm(timeOfDay)
``` 

```{r}
ggplot(full[rows.train,], aes(x = Month, fill = OutcomeType)) +
  geom_bar(position = 'fill', colour = 'black')

ggplot(full[rows.train,], aes(x = Weekday, fill = OutcomeType)) +
  geom_bar(position = 'fill', colour = 'black')
```

As I suspected, there is certain seasonal pattern: per months, july and december are good months for adoptions 
(matching end school and christmas holidays), and april worst month for adoptions.
Shelters are open 7 days a week, as result, saturdays and sundays are best days for adoptions. 
These are very reasonable deductions, following human customs.

### AnimalType

```{r}
outcomes <- full[rows.train,] %>% group_by(AnimalType, OutcomeType) %>% summarise(count = n())
ggplot(outcomes, aes(x = AnimalType, y = count, fill = OutcomeType)) + 
    geom_bar(stat = "identity", position = "fill", color = "black") +
    coord_flip() +
    labs(y = "Animals %", x = "Animal", title="Outcomes: cats & dogs")
```

Most visible difference is that dogs are more prone to be returned to the owner, rather cats are transferred to another shelter.

### Age

The feature `AgeuponOutcome` is a regular text field. But the real age of the animal could be a good predictor. 

```{r}
unique(full$AgeuponOutcome)
```

Applying certain rules, we are able to translate text field to numeric field. 
In the other side, precision is lost in the process (there is no difference between one year and one month and one year and 10 months); 
anyway, the lack of precission in inherent to the text field itself, the conversion didn't add any loss by itself.

```{r}
rows.fix.age <- is.na(full$AgeuponOutcome)
# fix missing age
full$Age.Value <- sapply(full$AgeuponOutcome, function(x) as.numeric(strsplit(x, split = ' ')[[1]][1]) )
full$Age.Unit <- sapply(full$AgeuponOutcome, function(x) gsub('s', '', strsplit(x, split = ' ')[[1]][2]))
convertToDays <- ifelse(full$Age.Unit == 'day', 1,
              ifelse(full$Age.Unit == 'week', 7,
              ifelse(full$Age.Unit == 'month', 30, 
              ifelse(full$Age.Unit == 'year', 365, NA))))
full$Age.Days <- full$Age.Value * convertToDays

full$Age.IsAdult <- full$Age.Days >= 365

summary(full$Age.Days)
``` 

Basically, text field is translated to a numeric field (using day units)

### Sex

```{r}
# imputate missing sex
prop.table(table(full$Sex[full$AnimalType=="Cat"]))
prop.table(table(full$Sex[full$AnimalType=="Dog"]))

rows.fix.sex.unknown <- grepl("Unknown", full$SexuponOutcome)
full$SexuponOutcome[rows.fix.sex.unknown] <- NA_character_
rows.fix.sex.na <- is.na(full$SexuponOutcome)

unique(full$SexuponOutcome)

full$SexuponOutcome <- as.factor(full$SexuponOutcome)
```

```{r mice, cache=T}
full.mice <- complete(mice(full[, c("AnimalType", "SexuponOutcome", "Breed", "Color", "Age.Days", "Breed.IsMix", "Breed.Main")], method = "rf"))
```

```{r}
#full$SexuponOutcome[rows.fix.sexoutcome] <- full$Sex[rows.fix.sexoutcome]
full$Sex.IsUnknown <- rows.fix.sex.unknown
full$Sex.IsIntact <- grepl("Intact", full$SexuponOutcome)
full$Sex <- gsub("Neutered |Intact |Spayed ", "", full$SexuponOutcome)
``` 

### Color

I think color is an important field. But right now, color field is text only, and we could squeeze more data from it.
Let's check a sample of colors 

```{r}
sample(unique(full$Color), 30)
```

There are some colors expressed as a mix of colors, and there are flat colors too. We can extract a feature (main color)
in order to identify first color in the field. Then, another feature can be used to know if the animal is single / mix colored.

```{r}
full$Color.Main <- sapply(full$Color, function(x) strsplit(x, split = '/| ')[[1]][1])
full$Color.IsFlat <- !grepl("/", full$Color)
``` 

There are several ways to encode colors. Most common encoding used in computers is rgb (red, green, blue components). 
But for humans, this encoding doesn't give any insight, this is the reason there are other encodings like 
HSV (hue, saturation, vibrance components) which is based in how humans perceive the colors.

R also has a builtin library for color names. Not all the colors in the `Color.Main` column are included in the library, 
but is a good starting point. From there, I will add some rules in order to complete the set of colors.

```{r}
convertToHsv <- function(colorname) {
    colorname <- tolower(colorname)
    rgb <- switch(colorname,
                  calico = "white",
                  cream = "#FFFDD0",
                  torbie = "black",
                  tortie = "black",
                  buff = "#F0DC82",
                  tricolor = "white",
                  seal = "#59260B",
                  fawn = "#E5AA70",
                  flame = "orangered",
                  sable = "saddlebrown",
                  liver = "sandybrown",
                  apricot = "#FBCEB1",
                  lynx = "antiquewhite2",
                  lilac = "#C8A2C8",
                  agouti = "ivory3",
                  ruddy = "indianred1",
                  silver = "#C0C0C0")
    
    if (is.null(rgb))
        rgb <- colorname
    
    return (rgb2hsv(col2rgb(rgb)))
}

colorsHSV <- full %>% 
    select(Color.Main) %>% 
    distinct() %>% 
    rowwise() %>% 
    mutate(Color.H = convertToHsv(Color.Main)["h", 1],
           Color.S = convertToHsv(Color.Main)["s", 1],
           Color.V = convertToHsv(Color.Main)["v", 1])

full <- full %>% inner_join(colorsHSV)
``` 

#### safe check

Look for examples of animaltype/breed and adopted/

## Processing

```{r}
factorVars <- c("OutcomeType", "AnimalType", "SexuponOutcome", "Breed", "Color", "Age.Unit", "Weekday", "Month", "Hour", "Breed.Main", "Color.Main", "Sex", "Breed.Hair")
full[factorVars] <- lapply(full[factorVars], function(x) as.factor(x))
``` 

```{r}

``` 

```{r}

``` 


## Notes
```{r}
save.image(file="shelter.rmd.RData")
sessionInfo()
```

[Dog Size Classification by Breed](http://www.petsatlanta.com/PageGeneral/breeds.html)
[Cat Classification By Breed](http://www.wikiwand.com/en/List_of_cat_breeds)
pre-process dog size
```{r eval=F}
dogSize <- read_csv(file = paste0(csv.folder, "dog_size.csv"))
dogBreedMatch <- str_match(dogSize$Breed, "(([a-zA-Z]*) - )?([ a-zA-Z]*)")
dogBreedMain <- paste(dogBreedMatch[, 4], ifelse(is.na(dogBreedMatch[, 3]), "", dogBreedMatch[, 3]))
dogBreedMain <- str_trim(dogBreedMain)
dogSize$Breed.Main <- dogBreedMain
#write_csv(dogSize, "dog_size.csv")
``` 

pre-process cat breeds
```{r}
catBreed <- read_csv(file = paste0(csv.folder, "cat_breed.csv"))
str(catBreed)
```
