---
title: "Shelter Journal"
date: "`r Sys.Date()`"
author: "[Fran Pérez](https://es.linkedin.com/in/franjperez)"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: journal
    highlight: tango
---

<style type="text/css">
div.sourceCode + table.table {
    width: 100%;
}

table.table {
    width: auto;
    margin: 20px auto;
    background-color: white;
}

.section.level5.collapse {
	padding: 10px;
	background-color: #FBE8E7;
    margin: 0 -10.5px;
    line-height: 1.42;
    border: 1px solid #cccccc;
    border-radius: 4px;
}

pre.r { 
    margin:0; 
}

pre.r  + pre {
    border-top: 0;
    margin: 0;
}

pre + p {
    margin-top: 10px;
}

blockquote {
    font-size: 16.5px !important;
}

div.sourceCode + p {
    margin-top: 10px;
}

pre.sourceCode.r {
    font-size: 13px;
}

p img {
    display: block;
    margin: 2px auto;
}

table + .caption, p + .caption {
    text-align: center;
    margin-top: -24px;
    padding: 10px;
    font-style: italic;
}

.glyphicon-eye-open {
    color: #eb6864;
    padding: 17px;
    padding-right: 0;
}

.level5 > h5  { display: none; }

</style>

## Introduction

[Shelter animal outcomes](https://www.kaggle.com/c/shelter-animal-outcomes)

[Quick & Dirty RandomForest](https://www.kaggle.com/mrisdal/shelter-animal-outcomes/quick-dirty-randomforest)

[XGB and 'Quick and Dirty' Ensemble](https://www.kaggle.com/jwjohnson314/shelter-animal-outcomes/xgb-and-quick-and-dirty-ensemble)

[dog groups](https://www.kaggle.com/andraszsom/shelter-animal-outcomes/dog-breeds-dog-groups)

[Uncertainty estimates of outcome types](https://www.kaggle.com/andraszsom/shelter-animal-outcomes/uncertainty-estimates-of-outcome-types)


## Feature engineering

Loading packages used for feature engineering

```{r warning=F,error=F,message=F}
library(readr) # csv file read / write
library(ggplot2) # visualization
library(ggthemes) # visualization
library(dplyr) # data manipulation
library(lubridate) # dates
library(caret) # classification algorithm
library(stringr)
library(mice) # imputation
library(stringdist) # string distance methods
library(knitr) # kable
``` 

```{r, include=F}
options(width = 120)
options(error = function() {
    save.image(file = "error.rmd.RData")
})
knitr::opts_chunk$set(fig.align = 'center')
set.seed(3874)
``` 

I used `read_csv`function in order to read training and test shelter data

```{r}
csv.folder <- ""
#kaggle folder
#csv.folder <- "../input/"
train <- read_csv(file = paste0(csv.folder, "train.csv"))
test <- read_csv(file = paste0(csv.folder, "test.csv"))
``` 

Then, I combine both data frames into a single one, named `full`. 
I did some column renaming in order to normalize column naming between datasets.

```{r}
full <- train %>% 
	rename(Id = AnimalID) %>% 
	select(-OutcomeSubtype) %>% 
	bind_rows(test %>% 
		rename(Id = ID) %>% 
		mutate(OutcomeType = NA, Id = as.character(Id))
	)

full.original <- full
rows.train <- !is.na(full$OutcomeType)

glimpse(full, width=100)
``` 

The full dataset contains `r dim(full)[2]` variables and `r dim(full)[1]` rows of data.

| Column     | Description    |
|------------|----------------|
| Id | animal Id |
| Name | animal name (could be nameless)|
| DataTime | timestamp |
| OutcomeType | animal outcome |
| AnimalType | dog or cat |
| SexuponOutcome | animal sex upon outcome |
| AgeuponOutcome | animal age upon outcome |
| Breed | animal's breed |
| Color | animal fur color |

There were another columns in the original files, but were not relevant for analysis.

### Breed

There are no missing values for breed feature. Let's review the data inside the column:

```{r}
# missing values should be zero
sum(is.na(full$Breed))
sample(unique(full$Breed), 30)
``` 

There are clearly two kinds of breeds: pure and mixed (in case of mixed breed, a "Mix" suffix appears on `Breed`)
```{r}
full$Breed.IsMix <- grepl("Mix$", full$Breed)
```

On a second look, there are some breeds with embedded hair features. This can be easily inspected using a regular expression looking for _hair_:
```{r}
unique(str_match(full$Breed,"(?i)[ a-z]*hair[a-z]*")[,1])
``` 

I count 5 different hair categories: long, medium, short, wire, hairless. Following this pattern, I create a new column named `Breed.Hair`

```{r}
full$Breed.Hair <-	ifelse (grepl("Longhair", full$Breed), "longhair",
					ifelse (grepl("Shorthair", full$Breed), "shorthair",
					ifelse (grepl("Wirehaired|Wire Hair|Wirehair", full$Breed), "wirehair",
					ifelse (grepl("Hairless", full$Breed), "hairless",
					ifelse (grepl("Medium Hair", full$Breed), "medium", "unknown")))))

table(full$Breed.Hair, full$AnimalType)
``` 

I will create a new column named `Breed.Main` which does not contained any 
_mix_ or _hair_ word, only the reference to the main breed.There are other suffixes used like "Dog" which could be removed, but for now, 
I think this is enough.

```{r}
# take first part of breed
breed.main <- sapply(full$Breed, function(x) gsub(' Mix', '', strsplit(x, split = '/')[[1]][1]))
# then remove hair
breed.main <- str_trim(gsub("Longhair|Shorthair|Wirehaired|Wire Hair|Wirehair|Hairless|Medium Hair", "", breed.main))

full$Breed.Main <- breed.main
``` 

If the breed is not mixed, we can foresee the average size of the animal. 
In the case of dogs, I will use a second dataframe, which correlates breeds with their size.

```{r}
# local / remote loading
dogSize <- read_csv("dog_size.csv")

rows.dog.not.mixed <- full$Breed.IsMix==F & full$AnimalType == "Dog"
```

<span class="glyphicon glyphicon-eye-open"></span> <a href="#dogSize_ss" class="button" data-toggle="collapse">safe check</a>

##### {#dogSize_ss .collapse}

The matching between breeds is not exact. But we can improve this matching using algorithms based on string distance.
I will check some of them:

```{r }

dogs.breeds <- full[rows.dog.not.mixed,c("Breed","Breed.Main")]
dogs.breeds.match <- amatch(dogs.breeds$Breed.Main, dogSize$Breed.Main, method = "jw", p = 0)
dogs.breeds$Breed.Match <- sapply(dogs.breeds.match, FUN = function (x) { ifelse(is.na(x), "UNK", dogSize$Breed.Main[x]) })
table(dogs.breeds$Breed.Match == "UNK")
#dogs.breeds[dogs.breeds$Breed.Match == "UNK",]

kable(unique(dogs.breeds[dogs.breeds$Breed.Main != dogs.breeds$Breed.Match & dogs.breeds$Breed.Match != "UNK",]))

dogs.breeds.match <- amatch(dogs.breeds$Breed.Main, dogSize$Breed.Main, method="osa")
dogs.breeds$Breed.Match <- sapply(dogs.breeds.match, FUN = function (x) { ifelse(is.na(x), "UNK", dogSize$Breed.Main[x]) })
table(dogs.breeds$Breed.Match == "UNK")

kable(unique(dogs.breeds[dogs.breeds$Breed.Main != dogs.breeds$Breed.Match & dogs.breeds$Breed.Match != "UNK",]))

dogs.breeds.match <- amatch(dogs.breeds$Breed.Main, dogSize$Breed.Main, method="cosine")
dogs.breeds$Breed.Match <- sapply(dogs.breeds.match, FUN = function (x) { ifelse(is.na(x), "UNK", dogSize$Breed.Main[x]) })
table(dogs.breeds$Breed.Match == "UNK")

kable(unique(dogs.breeds[dogs.breeds$Breed.Main != dogs.breeds$Breed.Match & dogs.breeds$Breed.Match != "UNK",]))
``` 

Best one seems to be `cosine`.

<a href="#dogSize_ss" data-toggle="collapse" class="text-right btn-block glyphicon glyphicon-remove"></a>

##### 


```{r}
full$Size[rows.dog.not.mixed] <- sapply(dogs.breeds.match, FUN = function (x) { ifelse(is.na(x), "UNK", dogSize$Size[x]) })
```

```{r include=F}
rm(dogSize)
```

### Name

Does having a name is correlated with the outcome?

```{r}
full$Name.NoName <- is.na(full$Name)
prop.table(table(full$Name.NoName))
``` 

### Date

We can extract several features from outcome date. 
This features are important because they can highlight seasonal behaviors.

```{r}
full$Hour <- hour(full$DateTime)
full$Weekday <- wday(full$DateTime, label = T)
full$Month <- month(full$DateTime, label = T)
full$Year <- year(full$DateTime)

timeOfDay <- ifelse(full$Hour > 5 & full$Hour < 11, 'morning',
                  ifelse(full$Hour > 10 & full$Hour < 16, 'midday',
                  ifelse(full$Hour > 15 & full$Hour < 20, 'lateday', 'night')))

full$TimeOfDay <- factor(timeOfDay, levels = c("morning", "midday", "lateday", "night"))
rm(timeOfDay)
``` 

```{r}
ggplot(full[rows.train,], aes(x = Month, fill = OutcomeType)) +
  geom_bar(position = 'fill', colour = 'black')

ggplot(full[rows.train,], aes(x = Weekday, fill = OutcomeType)) +
  geom_bar(position = 'fill', colour = 'black')
```

As I suspected, there is certain seasonal pattern: per months, july and december are good months for adoptions 
(matching end school and christmas holidays), and april worst month for adoptions.
Shelters are open 7 days a week, as result, saturdays and sundays are best days for adoptions. 
These are very reasonable deductions, following human customs.

### AnimalType

```{r}
outcomes <- full[rows.train,] %>% group_by(AnimalType, OutcomeType) %>% summarise(count = n())
ggplot(outcomes, aes(x = AnimalType, y = count, fill = OutcomeType)) + 
    geom_bar(stat = "identity", position = "fill", color = "black") +
    coord_flip() +
    labs(y = "Animals %", x = "Animal", title="Outcomes: cats & dogs")
```

Most visible difference is that dogs are more prone to be returned to the owner, rather cats are transferred to another shelter.

### Age

The feature `AgeuponOutcome` is a regular text field. But the real age of the animal could be a good predictor. 

```{r}
unique(full$AgeuponOutcome)
```

Applying certain rules, we are able to translate text field to numeric field. 
In the other side, precision is lost in the process (there is no difference between one year and one month and one year and 10 months); 
anyway, the lack of precission in inherent to the text field itself, the conversion didn't add any loss by itself.

```{r}
rows.fix.age <- is.na(full$AgeuponOutcome)
# fix missing age
full$Age.Value <- sapply(full$AgeuponOutcome, function(x) as.numeric(strsplit(x, split = ' ')[[1]][1]) )
full$Age.Unit <- sapply(full$AgeuponOutcome, function(x) gsub('s', '', strsplit(x, split = ' ')[[1]][2]))
convertToDays <- ifelse(full$Age.Unit == 'day', 1,
              ifelse(full$Age.Unit == 'week', 7,
              ifelse(full$Age.Unit == 'month', 30, 
              ifelse(full$Age.Unit == 'year', 365, NA))))
full$Age.Days <- full$Age.Value * convertToDays

full$Age.IsAdult <- full$Age.Days >= 365

summary(full$Age.Days)
``` 

Basically, text field is translated to a numeric field (using day units)

### Sex

```{r}
# imputate missing sex
prop.table(table(full$Sex[full$AnimalType=="Cat"]))
prop.table(table(full$Sex[full$AnimalType=="Dog"]))

rows.fix.sex.unknown <- grepl("Unknown", full$SexuponOutcome)
full$SexuponOutcome[rows.fix.sex.unknown] <- NA_character_
rows.fix.sex.na <- is.na(full$SexuponOutcome)
table(rows.fix.sex.na)

unique(full$SexuponOutcome)

#full$SexuponOutcome <- as.factor(full$SexuponOutcome)
```

```{r mice, cache=T}
full.mice <- complete(mice(full[, c("AnimalType", "SexuponOutcome", "Breed", "Color", "Age.Days", "Breed.IsMix", "Breed.Main")], method = "rf"))
```

<span class="glyphicon glyphicon-eye-open"></span> <a href="#mice_ss" class="button" data-toggle="collapse">safe check</a>

##### {#mice_ss .collapse}


```{r}
#new ages
kable(full.mice[rows.fix.age,])

#new sex .. there are too many!!
#kable(full.mice[rows.fix.sex.na,])

age.type <- full[!rows.fix.age,] %>% 
	mutate (AgeType = "Age (original)") %>% 
	select (Age.Days, AgeType) %>% 
	bind_rows(full.mice %>% 
				mutate (AgeType="Age (after MICE)") %>% 
				select (Age.Days, AgeType))
ggplot(age.type, aes(x = Age.Days)) +
	geom_density(aes(fill = factor(AgeType)), alpha = 0.6) +
	#scale_fill_manual(values = c(alpha("green", 0.25), alpha("blue", 0.4))) +
	labs(x = 'Age', fill = '')
```

<a href="#mice_ss" data-toggle="collapse" class="text-right btn-block glyphicon glyphicon-remove"></a>

##### 


```{r}
#full$SexuponOutcome[rows.fix.sexoutcome] <- full$Sex[rows.fix.sexoutcome]
full$Age.Days <- full.mice$Age.Days
full$SexuponOutcome <- full.mice$SexuponOutcome
full$Sex.IsUnknown <- rows.fix.sex.unknown
full$Sex.IsIntact <- grepl("Intact", full$SexuponOutcome)
full$Sex <- gsub("Neutered |Intact |Spayed ", "", full$SexuponOutcome)
``` 

```{r include=F}
rm (age.type, full.mice)
```

### Color

I think color is an important field. But right now, color field is text only, and we could squeeze more data from it.
Let's check a sample of colors 

```{r}
sample(unique(full$Color), 30)
```

There are some colors expressed as a mix of colors, and there are flat colors too. We can extract a feature (main color)
in order to identify first color in the field. Then, another feature can be used to know if the animal is single / mix colored.

```{r}
full$Color.Main <- sapply(full$Color, function(x) strsplit(x, split = '/| ')[[1]][1])
full$Color.IsFlat <- !grepl("/", full$Color)
``` 

There are several ways to encode colors. Most common encoding used in computers is rgb (red, green, blue components). 
But for humans, this encoding doesn't give any insight, this is the reason there are other encodings like 
HSV (hue, saturation, vibrance components) which is based in how humans perceive the colors.

R also has a builtin library for color names. Not all the colors in the `Color.Main` column are included in the library, 
but is a good starting point. From there, I will add some rules in order to complete the set of colors.

```{r}
convertToHsv <- function(colorname) {
    colorname <- tolower(colorname)
    rgb <- switch(colorname,
                  calico = "white",
                  cream = "#FFFDD0",
                  torbie = "black",
                  tortie = "black",
                  buff = "#F0DC82",
                  tricolor = "white",
                  seal = "#59260B",
                  fawn = "#E5AA70",
                  flame = "orangered",
                  sable = "saddlebrown",
                  liver = "sandybrown",
                  apricot = "#FBCEB1",
                  lynx = "antiquewhite2",
                  lilac = "#C8A2C8",
                  agouti = "ivory3",
                  ruddy = "indianred1",
                  silver = "#C0C0C0")
    
    if (is.null(rgb))
        rgb <- colorname
    
    return (rgb2hsv(col2rgb(rgb)))
}

colorsHSV <- full %>% 
    select(Color.Main) %>% 
    distinct() %>% 
    rowwise() %>% 
    mutate(Color.H = convertToHsv(Color.Main)["h", 1],
           Color.S = convertToHsv(Color.Main)["s", 1],
           Color.V = convertToHsv(Color.Main)["v", 1])

full <- full %>% inner_join(colorsHSV)
``` 

#### safe check

Look for examples of animaltype/breed and adopted/

## Processing

```{r}
factorVars <- c("OutcomeType", "AnimalType", "SexuponOutcome", "Breed", "Color", "Age.Unit", "Weekday", "Month", "Hour", "Breed.Main", "Color.Main", "Sex", "Breed.Hair")
full[factorVars] <- lapply(full[factorVars], function(x) as.factor(x))

glimpse(full, width=100)
``` 

```{r}

``` 

```{r}

``` 


## Notes
```{r}
save.image(file="shelter.rmd.RData")
sessionInfo()
```

[Dog Size Classification by Breed](http://www.petsatlanta.com/PageGeneral/breeds.html)
[Cat Classification By Breed](http://www.wikiwand.com/en/List_of_cat_breeds)
pre-process dog size
```{r eval=F}
dogSize <- read_csv(file = paste0(csv.folder, "dog_size.csv"))
#dogBreedMatch <- str_match(dogSize$Breed, "(([a-zA-Z]*) - )?([ a-zA-Z]*)")
#dogBreedMain <- paste(dogBreedMatch[, 4], ifelse(is.na(dogBreedMatch[, 3]), "", dogBreedMatch[, 3]))
dogBreedMain <- str_trim(gsub("Longhair|Shorthair|Wirehaired|Wire Hair|Wirehair|Hairless|Medium Hair", "", dogBreedMain))
dogSize$Breed.Main <- dogBreedMain
#write_csv(dogSize, "dog_size.csv")
``` 

pre-process cat breeds
```{r}
#catBreed <- read_csv(file = paste0(csv.folder, "cat_breed.csv"))
#str(catBreed)
```
